---
title: "R for Spatial Analysis"
author: "Aaron Kleyn"
date: "May 17, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r, load income_grid, include=FALSE}
load("C:/Users/Aaron/Downloads/income_grid.rda")
```
```{r, load land_cover, include=FALSE}
load("C:/Users/Aaron/Downloads/land_cover.rda")
```
```{r, load migration, include=FALSE}
load("C:/Users/Aaron/Downloads/migration.rda")
```
```{r, load neighborhoods, include=FALSE}
load("C:/Users/Aaron/Downloads/neighborhoods.rda")
```
```{r, load nyc_income, include=FALSE}
load("C:/Users/Aaron/Downloads/nyc_income.rda")
```
```{r, load pop, include=FALSE}
load("C:/Users/Aaron/Downloads/pop.rda")
```
```{r, load prop_by_age, include=FALSE}
load("C:/Users/Aaron/Downloads/prop_by_age.rda")
```
```{r, load pop_by_age, include=FALSE}
load("C:/Users/Aaron/Downloads/pop_by_age.rda")
```
```{r, load preds, include=FALSE}
load("C:/Users/Aaron/Downloads/preds.rda")

```
```{r, load water, include = FALSE}
load("C:/Users/Aaron/Downloads/water.rda")
```


The following three plots will be used to construct a demographic plot New York City and the surrounding regions. 

```{r, warning = FALSE}
library(ggplot2)
library(sp)
library(RColorBrewer)
library(raster)
library(tmap)


# 9 steps on the RColorBrewer "BuPu" palette: blups
blups <- brewer.pal(9, "BuPu")

# Add scale_fill_gradientn() with the blups palette
ggplot(preds) +
  geom_tile(aes(lon, lat, fill = predicted_price), alpha = 0.8) +
  scale_fill_gradientn(colors = blups)

library(viridisLite)
# viridisLite viridis palette with 9 steps: vir
vir <- viridis(9)

# Add scale_fill_gradientn() with the vir palette
ggplot(preds) +
  geom_tile(aes(lon, lat, fill = predicted_price), alpha = 0.8) +
  scale_fill_gradientn(colors = vir)
  
# mag: a viridisLite magma palette with 9 steps
mag <- viridisLite::magma(7)

```



```{r, warning = FALSE }
library(classInt)
library(RColorBrewer)

# Create 5 "pretty" breaks with classIntervals()
classIntervals(values(prop_by_age[["age_18_24"]]), 
               n = 5, style = "pretty")

# Create 5 "quantile" breaks with classIntervals()
classIntervals(values(prop_by_age[["age_18_24"]]), 
               n = 5, style = "quantile")

# Use 5 "quantile" breaks in tm_raster()
tm_shape(prop_by_age) +
  tm_raster("age_18_24", palette = mag, style = "quantile") +
  tm_legend(position = c("right", "bottom"))

# Create histogram of proportions
hist(values(prop_by_age)[, "age_18_24"])

# Use fixed breaks in tm_raster()
tm_shape(prop_by_age) +
  tm_raster("age_18_24", palette = mag,
    style = "fixed", breaks = c(0.025, 0.05, 0.1, 0.2, 0.25, 0.3, 1))

# Save your plot to "prop_18-24.html"
save_tmap(filename = "prop_18-24.html")

```

The following code will derive a plot of net migration in the US from 1990 - 2000

```{r}
# Print migration

migration

# Diverging "RdGy" palette
red_gray <- brewer.pal(7, "RdGy")

# Use red_gray as the palette 
tm_shape(migration) +
  tm_raster(palette = red_gray) +
  tm_legend(outside = TRUE, outside.position = c("bottom"))

# Add fixed breaks 
tm_shape(migration) +
  tm_raster(palette = red_gray, style = "fixed", 
     breaks = c(-5e6, -5e3, -5e2, -5e1, 5e1, 5e2, 5e3, 5e6)) +
  tm_legend(outside = TRUE, outside.position = c("bottom"))

```


The following R code will derive a plot of vegetation/climate cover based on longitudinal and latitudinal coordinates. 
```{r}
library(raster)

# Plot land_cover
tm_shape(land_cover) +
  tm_raster() 

# Palette like the ggplot2 default
hcl_cols <- hcl(h = seq(15, 375, length = 9), 
                c = 100, l = 65)[-9]

# Use hcl_cols as the palette
tm_shape(land_cover) +
  tm_raster(palette = hcl_cols) 

# Examine levels of land_cover
levels(land_cover)

# A set of intuitive colors
intuitive_cols <- c(
  "darkgreen",
  "darkolivegreen4",
  "goldenrod2",
  "seagreen",
  "wheat",
  "slategrey",
  "white",
  "lightskyblue1"
)

# Use intuitive_cols as palette
tm_shape(land_cover) +
  tm_raster(palette = intuitive_cols) +
  tm_legend(position = c("left", "bottom"))

```

The following R code will be used to develop a plot of average annual income in Manhattan. 

```{r, eval = FALSE}
library(sp)
library(rgdal)

# Use dir() to find directory name
dir()

# Call dir() with directory name
dir('nynta_16c')

# Read in shapefile with readOGR(): neighborhoods
neighborhoods <- readOGR('nynta_16c', 'nynta')
```

```{r}
# summary() of neighborhoods
summary(neighborhoods)

# Plot neighboorhoods
plot(neighborhoods)

```


```{r, eval = FALSE}
library(raster) 

# Call dir()
dir()

# Call dir() on the directory
dir('nyc_grid_data')

# Use raster() with file path: income_grid
income_grid <- raster('nyc_grid_data/m5602ahhi00.tif')
```

```{r}
# Call summary() on income_grid
summary(income_grid)

# Call plot() on income_grid
plot(income_grid)

```



```{r, message = FALSE, warning = FALSE}
library(sp)
library(tigris)
# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = 'NY', county = 'New York', cb = TRUE)

# proj4string() on nyc_tracts and neighborhoods
proj4string(nyc_tracts)
proj4string(neighborhoods)

# coordinates() on nyc_tracts and neighborhoods
head(coordinates(nyc_tracts))
head(coordinates(neighborhoods))

# plot() neighborhoods and nyc_tracts
plot(neighborhoods)
plot(nyc_tracts, add = TRUE, col = 'red')

```


```{r}
library(sp)
library(raster)
library(tigris)
# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = 'NY', county = 'New York', cb = TRUE)
# Use spTransform on neighborhoods: neighborhoods
neighborhoods <- spTransform(neighborhoods,   
                             proj4string(nyc_tracts))

# head() on coordinates() of neighborhoods
head(coordinates(neighborhoods))

# Plot neighborhoods, nyc_tracts and water
plot(neighborhoods)
plot(nyc_tracts, add = TRUE, col = "red")
plot(water, add = TRUE, col = "blue")

```

```{r}
library(sp)
library(tigris)
# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = 'NY', county = 'New York', cb = TRUE)
# Use str() on nyc_income and nyc_tracts@data
str(nyc_income)
str(nyc_tracts@data)

# Highlight tract 002201 in nyc_tracts
plot(nyc_tracts)
plot(nyc_tracts[nyc_tracts$TRACTCE == "002201", ], 
     col = "red", add = TRUE)

# Set nyc_tracts@data to nyc_income
nyc_tracts@data <- nyc_income

# Highlight tract 002201 in nyc_tracts
plot(nyc_tracts)
plot(nyc_tracts[nyc_tracts$tract == "002201", ], 
     col = "red", add = TRUE)

```


```{r}
library(tigris)
# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = 'NY', county = 'New York', cb = TRUE)
# Check for duplicates in nyc_income
any(duplicated(nyc_income$tract))

# Check for duplicates in nyc_tracts
any(duplicated(nyc_tracts$TRACTCE))

# Check nyc_tracts in nyc_income
all((nyc_tracts$TRACTCE %in% nyc_income$tract))

# Check nyc_income in nyc_tracts
all((nyc_income$tract %in% nyc_tracts$TRACTCE))

```


```{r}
library(sp)
library(tmap)
library(tigris)
# Call tracts(): nyc_tracts
nyc_tracts <- tracts(state = 'NY', county = 'New York', cb = TRUE)
# Merge nyc_tracts and nyc_income: nyc_tracts_merge
nyc_tracts_merge <- sp::merge(nyc_tracts, nyc_income, by.x = "TRACTCE", by.y = "tract")

# Call summary() on nyc_tracts_merge
summary(nyc_tracts_merge)

# Choropleth with col mapped to estimate
tm_shape(nyc_tracts_merge) +
  tm_fill(col = "estimate") 

```


```{r}
library(tmap)
library(tigris)
tm_shape(nyc_tracts_merge) +
  tm_fill(col = "estimate") +
  # Add a water layer, tm_fill() with col = "grey90"
  tm_shape(water) +
  tm_fill(col = "grey90") +
  # Add a neighborhood layer, tm_borders()
  tm_shape(neighborhoods) +
  tm_borders() 

```

```{r}
library(tmap)

# Find unique() nyc_tracts_merge$COUNTYFP
unique(nyc_tracts_merge$COUNTYFP)

# Add logical expression to pull out New York County
manhat_hoods <- neighborhoods[neighborhoods$CountyFIPS == "061", ]

tm_shape(nyc_tracts_merge) +
  tm_fill(col = "estimate") +
  tm_shape(water) +
  tm_fill(col = "grey90") +
  # Edit to use manhat_hoods instead
  tm_shape(manhat_hoods) +
  tm_borders() +
  # Add a tm_text() layer
  tm_text(text = "NTAName")

```


```{r}
library(tmap)

# gsub() to replace " " with "\n"
manhat_hoods$name <- gsub(" ", "\n", manhat_hoods$NTAName)

# gsub() to replace "-" with "/\n"
manhat_hoods$name <- gsub("-", "/\n", manhat_hoods$name)

# Edit to map text to name, set size to 0.5
tm_shape(nyc_tracts_merge) +
    tm_fill(col = "estimate") +
  tm_shape(water) +
    tm_fill(col = "grey90") +
  tm_shape(manhat_hoods) +
    tm_borders() +
    tm_text(text = "name", size = 0.5)

```


```{r}
library(tmap)

tm_shape(nyc_tracts_merge) +
  # Add title and change palette
  tm_fill(col = "estimate", 
          title = "Median Income",
          palette = "Greens") +
  # Add tm_borders()
  tm_borders(col = "grey60", lwd = 0.5) +
  tm_shape(water) +
  tm_fill(col = "grey90") +
  tm_shape(manhat_hoods) +
  # Change col and lwd of neighborhood boundaries
  tm_borders(col = "grey40", lwd = 2) +
  tm_text(text = "name", size = 0.5) +
  # Add tm_credits()
  tm_credits("Source: ACS 2014 5-year Estimates, \n accessed via acs package", 
             position = c("right", "bottom"))
```


```{r, eval = FALSE}
# Save map as "nyc_income_map.png"
save_tmap(filename = "nyc_income_map.png", width = 4, height = 7)

```

